Type cones of $\b{g}$-vector fans

### 2019-04-02 thirdPaper/draft/typeConeAssociahedra.tex
mesh mutations

\begin{definition}
The \defn{$\b{c}$-vector} of a cluster variable $x$ in a seed~$\seed$ of~$\principalClusterAlgebra$ is the vector
\(
\cvector{\seed}{x} = \cvectorFull{\B_\circ}{\seed}{x} \eqdef \sum_{y \in \cluster_\circ} c_{yx} \, \simpleRoot_y
\)
of exponents of~$p_x = \prod_{y \in \cluster_\circ} (p_y)^{c_{yx}}$.
We denote by~$\cvectors{\seed} = \cvectorsFull{\B_\circ}{\seed} \eqdef \set{\cvectorFull{\B_\circ}{\seed}{x}}{x \in \seed}$ the set of $\b{c}$-vectors of a seed~$\seed$.
\end{definition}

@@

\begin{definition}
%Consider a seed~$\seed = (\B, \cluster)$ and a cluster variable~$x \in \cluster$.
The mutation of a seed~$\seed = (\B, \cluster)$ in the direction of a cluster variable~$x \in \cluster$ is a \defn{mesh mutation} that \defn{starts} (resp.~\defn{ends}) at~$x$ if the entries~$b_{xy}$ for~$y \in \cluster$ are all non-negative (resp.~all non-positive).
A mesh mutation is \defn{initial} if it starts at a cluster variable of an initial seed.
\end{definition}

We denote by~$\meshes$ the set of all pairs~$(x,x')$ where~$x$ and~$x'$ are two cluster variables of~$\principalClusterAlgebra$ which are exchangeable via a non-initial mesh mutation.

\begin{lemma}
\label{lem:bijectionMeshMutations}
There is precisely one non-initial mesh directed by each positive $\b{c}$-vector of the cluster algebra~$\principalClusterAlgebra$. Therefore~$|\principalVariables| = |\meshes| + n$.
\end{lemma}

@@

Our main result states that the linear dependences between the $\b{g}$-vectors of any mutation can be decomposed into positive combinations of linear dependences arising from meshes.

\begin{proposition}
\label{prop:extremalExchangeablePairsCA}
For any finite type exchange matrix~$\B_\circ$ (acyclic or not), the non-initial mesh mutations precisely correspond to the extremal exchangeable pairs of the cluster fan~$\gvectorFan[\B_\circ]$.
\end{proposition}

This statement will be proved in full generality in \cref{???}.
We proved it in \cref{prop:extremalExchangeablePairsAsso} for acyclic initial seeds in type~$A$. To provide an intuition of the result, we sketch here a proof for arbitrary initial seeds in type~$A$.

\begin{proof}[Sketch of \cref{prop:extremalExchangeablePairsCA} in type~$A$]
We need to prove that the normal vector~$\b{n}(\B_\circ, y, y')$ for any exchangeable pair~$\{y,y'\}$ is a positive linear combination of the vectors~$\b{n}(\B_\circ, x, x')$ for non-initial meshes.
\vincent{Do we really want to do that?}
\end{proof}

Our next statement follows from~\cref{prop:extremalExchangeablePairsCA,lem:bijectionMeshMutations}.

@@

Combining \cref{coro:simplicialTypeCone,coro:simplicialTypeConeCA,prop:extremalExchangeablePairsCA}, we derive the following description of all polytopal realizations of the cluster fan~$\gvectorFan[\B_\circ]$.
This result was stated in~\cite{BazierMatteDouvilleMousavandThomasYildirim} in the special situation of acyclic seeds in simply-laced types.
Note that our proof is quite different from~\cite{BazierMatteDouvilleMousavandThomasYildirim} as it relies on our type cone approach.

>>>>

\begin{definition}
The \defn{$\b{c}$-vector} of a cluster variable $x$ in a seed~$\seed$ of~$\principalClusterAlgebra$ is the vector
\(
{\cvector{\seed}{x} = \cvectorFull{\B_\circ}{\seed}{x} \eqdef \sum_{y \in \cluster_\circ} c_{yx} \, \simpleRoot_y}
\)
of exponents of~$p_x = \prod_{y \in \cluster_\circ} (p_y)^{c_{yx}}$.
We denote by ${\cvectors{\seed} = \cvectorsFull{\B_\circ}{\seed} \eqdef \set{\cvectorFull{\B_\circ}{\seed}{x}}{x \in \seed}}$ the set of $\b{c}$-vectors of a seed~$\seed$.
\end{definition}

@@

\begin{definition}
%Consider a seed~$\seed = (\B, \cluster)$ and a cluster variable~$x \in \cluster$.
The mutation of a seed~$\seed = (\B, \cluster)$ in the direction of a cluster variable~$x \in \cluster$ is a \defn{mesh mutation} that \defn{starts} (resp.~\defn{ends}) at~$x$ if the entries~$b_{xy}$ for~$y \in \cluster$ are all non-negative (resp.~all non-positive).
A mesh mutation is \defn{initial} if it starts at a cluster variable of an initial seed.
We denote by~$\meshes$ the set of all pairs~$(x,x')$ where~$x$ and~$x'$ are two cluster variables of~$\principalClusterAlgebra$ which are exchangeable via a non-initial mesh mutation.
\end{definition}

\begin{lemma}
\label{lem:bijectionMeshMutations}
There is precisely one non-initial mesh directed by each positive $\b{c}$-vector of the cluster algebra~$\principalClusterAlgebra[\B_\circ^\vee]$. Therefore~$|\principalVariables| = |\meshes| + n$.
\end{lemma}

@@

The following will be shown in \cref{sec:clusterCategories}.

\begin{proposition}
\label{prop:meshMutations}
For any finite type exchange matrix~$\B_\circ$ (acyclic or not, simply-laced or not), the linear dependence between the $\b{g}$-vectors of any mutation can be decomposed into positive combinations of linear dependences between $\b{g}$-vectors of non-initial mesh mutations.
\end{proposition}

We derive from \cref{prop:meshMutations} the following description of the facets of the type cone of the $\b{g}$-vector fan.

\begin{theorem}
\label{thm:extremalExchangeablePairsCA}
For any finite type exchange matrix~$\B_\circ$ (acyclic or not, simply-laced or not), the non-initial mesh mutations precisely correspond to the extremal exchangeable pairs of the cluster fan~$\gvectorFan[\B_\circ]$.
\end{theorem}

\begin{proof}
\end{proof}

%This statement will be proved in full generality in \cref{???}.
%We proved it in \cref{prop:extremalExchangeablePairsAsso} for acyclic initial seeds in type~$A$. To provide an intuition of the result, we sketch here a proof for arbitrary initial seeds in type~$A$.
%
%\begin{proof}[Sketch of \cref{thm:extremalExchangeablePairsCA} in type~$A$]
%We need to prove that the normal vector~$\b{n}(\B_\circ, y, y')$ for any exchangeable pair~$\{y,y'\}$ is a positive linear combination of the vectors~$\b{n}(\B_\circ, x, x')$ for non-initial meshes.
%\vincent{Do we really want to do that?}
%\end{proof}

We also obtain from \cref{thm:extremalExchangeablePairsCA} the following surprising property.

Our next statement follows from~\cref{thm:extremalExchangeablePairsCA,lem:bijectionMeshMutations}.

@@

Combining \cref{coro:simplicialTypeCone,coro:simplicialTypeConeCA,thm:extremalExchangeablePairsCA}, we derive the following description of all polytopal realizations of the cluster fan~$\gvectorFan[\B_\circ]$.
This result was stated in~\cite{BazierMatteDouvilleMousavandThomasYildirim} in the special situation of acyclic seeds in simply-laced types.
Note that our proof is quite different from~\cite{BazierMatteDouvilleMousavandThomasYildirim} as it relies on our type cone approach.
